Reference Tracking Strategies of Native and Late Signers in Turkish Sign Language

### 2022-08-06 revised_manuscript.Rmd
updated files

For more inference on the data and having pairwise comparisons, we fit a regression model to RE Type using discourse status and acquisition group as predictors. We used the brms package [@burkner2018] in R, with a categorical distribution on the responses and sum contrasts for the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late). As the overall number of produced pronouns was very little (40 data points), we excluded pronominal use from the following analyses. For the current analysis, we did not define priors ourselves and used the preset uninformed priors. Figure 2[^2] shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals. There are 3 responses we are considering in this model. They are; Classifier, Nominative, and Null. Classifier response is the level that other responses Nominative and Null are compared against.

@@

[^3]: As in other sign languages, in TİD, the lexical meaning of signs has an effect on the morphological grouping. For instance, usually (di)transitive agreement verbs denote some sort of physical or metaphorical transfer of an object or abstract notion (Meir 1999). As for plain verbs, both phonology of a sign and its semantics have an effect on morphology. For instance, \textsc{eat} in TİD is signed nearby the mouth and since it needs to remain in that phonological location, it is not suitable for agreement and thus it is regarded as a plain verb. Likewise the experiencer verb \textsc{love} is signed on the torso and as such it does not take part in agreement. These generalizations noted we should also be cautious in labelling a verb as categorially one type or the other. For instance, some intransitive plain verbs like \textsc{fall} or \textsc{jump} in TİD may display spatial person agreement with the single argument in the clause (Gökgöz and Sevgi, 2020). See also for a similar discussion in Brazilian Sign Language.}.

@@

```{r model-camera, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late)", fig.height=3.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
df_model_1 %<>% mutate(model = "All")
df_model_camera %<>% mutate(model = "Camera")
model_comparison <- rbind(df_model_1, df_model_camera)

p4 <- model_comparison %>% 
  ggplot(aes(m, y=factor(parameter, levels = rev(levels(factor(parameter)))), group = model, color = model)) +
  geom_point(size = 2.5, position = position_dodge(width = 0.6)) +
  geom_errorbarh(aes(xmin = l, xmax = h), alpha = 1, height = 0, size = 1.5, position = position_dodge(width = 0.6)) +
  geom_errorbarh(aes(xmin = ll, xmax = hh), height=0, position = position_dodge(width = 0.6)) + vline_0(linetype = "dashed") +
  xlab("Estimate (log odds)") + ylab("Coefficients") + theme_bw() + theme_apa() +
  scale_color_grey() +
  scale_fill_grey() 
p4
```

>>>>

For more inference on the data and having pairwise comparisons, we fit a regression model to RE Type using discourse status and acquisition group as predictors. We used the brms package [@burkner2018] in R, with a categorical distribution on the responses and sum contrasts for the predictors Discourse Status (Introduction, Maintenance, Re-introduction) and Participant Group (Native, Late). As the overall number of produced pronouns was very little (40 data points), we excluded pronominal use from the following analyses. For the current analysis, we did not define priors ourselves and used the preset uninformed priors. Figure 2[^2] shows the posterior probability distributions where the dot is the median estimate, the thick line is the 50% credible intervals, and the thin line is the 95% credible intervals. There are 3 responses we are considering in this model. They are Classifier, Nominative, and Null. Classifier response is the level that other responses Nominative and Null are compared against.

@@

[^3]: As in other sign languages, in TİD, the lexical meaning of signs has an effect on the morphological grouping. For instance, usually (di)transitive agreement verbs denote some sort of physical or metaphorical transfer of an object or abstract notion [@meir1999]. As for plain verbs, both phonology of a sign and its semantics have an effect on morphology. For instance, \textsc{eat} in TİD is signed nearby the mouth and since it needs to remain in that phonological location, it is not suitable for agreement and thus it is regarded as a plain verb. Likewise the experiencer verb \textsc{love} is signed on the torso and as such it does not take part in agreement. These generalizations noted we should also be cautious in labelling a verb as categorially one type or the other. For instance, some intransitive plain verbs like \textsc{fall} or \textsc{jump} in TİD may display spatial person agreement with the single argument in the clause [@gokgoz2020]. See @lourenço2018 also for a similar discussion in Brazilian Sign Language.}.

@@

```{r model-camera, fig.cap="Regression model plot for the RE Type (Nominal, Null) with the predictors Discourse Status (Introduction, Maintenance, Re-introduction), Participant Group (Native, Late) and Narration Style (all responses vs. only camera)", fig.height=3.5, fig.pos="hbt!", fig.width=4, echo=FALSE, message=FALSE, cache=FALSE, results='asis', warnings=FALSE}
df_model_1 %<>% mutate(model = "All")
df_model_camera %<>% mutate(model = "Camera")
model_comparison <- rbind(df_model_1, df_model_camera)

p4 <- model_comparison %>% 
  ggplot(aes(m, y=factor(parameter, levels = rev(levels(factor(parameter)))), group = model, color = model)) +
  geom_point(size = 2.5, position = position_dodge(width = 0.6)) +
  geom_errorbarh(aes(xmin = l, xmax = h), alpha = 1, height = 0, size = 1.5, position = position_dodge(width = 0.6)) +
  geom_errorbarh(aes(xmin = ll, xmax = hh), height=0, position = position_dodge(width = 0.6)) + vline_0(linetype = "dashed") +
  xlab("Estimate (log odds)") + ylab("Coefficients") + theme_bw() + theme_apa() +
  scale_color_grey() +
  scale_fill_grey() 
p4

```

*Note:* "All" indicates model results including all participants (both camera and experimenter) whereas "camera" indicates the results from only those who narrated to the camera. That is, the 4 signers who narrated to the experimenter were from analysis in the second dataset.

### 2022-08-06 revised_manuscript.tex
updated files

For more inference on the data and having pairwise comparisons, we fit a
regression model to RE Type using discourse status and acquisition group
as predictors. We used the brms package (Bürkner, 2018) in R, with a
categorical distribution on the responses and sum contrasts for the
predictors Discourse Status (Introduction, Maintenance, Re-introduction)
and Participant Group (Native, Late). As the overall number of produced
pronouns was very little (40 data points), we excluded pronominal use
from the following analyses. For the current analysis, we did not define
priors ourselves and used the preset uninformed priors. Figure
2\footnote{We ran another model where we used the Nativeness predictor
  with 3 levels 0-3:Native, 4-7:Late1, and 7-above:Late2. However, our
  model did not converge properly and the credible intervals were very
  wide spread, yielding the results uninformative. We provide the model
  results as a plot in \ref{section:d}.} shows the posterior probability
distributions where the dot is the median estimate, the thick line is
the 50\% credible intervals, and the thin line is the 95\% credible
intervals. There are 3 responses we are considering in this model. They
are; Classifier, Nominative, and Null. Classifier response is the level
that other responses Nominative and Null are compared against.

@@

As for the distribution of the null markers, approximately 90\% of the
plain verbs were used to maintain a previous referent, making these
constructions least implicit REs used for TİD\footnote{As in other sign
  languages, in TİD, the lexical meaning of signs has an effect on the
  morphological grouping. For instance, usually (di)transitive agreement
  verbs denote some sort of physical or metaphorical transfer of an
  object or abstract notion (Meir 1999). As for plain verbs, both
  phonology of a sign and its semantics have an effect on morphology.
  For instance, \textsc{eat} in TİD is signed nearby the mouth and since
  it needs to remain in that phonological location, it is not suitable
  for agreement and thus it is regarded as a plain verb. Likewise the
  experiencer verb \textsc{love} is signed on the torso and as such it
  does not take part in agreement. These generalizations noted we should
  also be cautious in labelling a verb as categorially one type or the
  other. For instance, some intransitive plain verbs like \textsc{fall}
  or \textsc{jump} in TİD may display spatial person agreement with the
  single argument in the clause (Gökgöz and Sevgi, 2020). See also for a
  similar discussion in Brazilian Sign Language.\}.}. Agreement verbs
and CA were the following implicit markers of reference, 0.77 and 0.72
of which were used for maintained contexts, respectively. This contrasts
with some of the research (Czubek, 2017) in which agreement verbs ranked
lowest on scale, indicating that these constructions were more implicit
compared to plain verbs. However, there seems to be a common ground for
the position of CA markers which have been claimed to rank higher than
other null markers as in the present study (Barberà and Massó, 2009;
Czubek, 2017; Frederiksen and Mayberry, 2016). In Figure
\ref{fig:scale}, we report the distribution of REs in TİD as a function
of referent accessibility and compare our accessibility hierarchy with
that of (Frederiksen and Mayberry, 2016) for ASL. In the hierarchies,
the left most REs (e.g., modified and bare nouns for TİD) are used when
the accessibility is low. If the referent is already active in the
addressee's mind and its accessibility is high, then the right most
markers (e.g., agreement and plain verbs for TİD) are selected.

@@

\begin{figure}
\centering
\includegraphics{revised_manuscript_files/figure-latex/model-camera-1.pdf}
\caption{Regression model plot for the RE Type (Nominal, Null) with the
predictors Discourse Status (Introduction, Maintenance, Re-introduction)
and Participant Group (Native, Late)}
\end{figure}

@@

@@

@@

>>>>

For more inference on the data and having pairwise comparisons, we fit a
regression model to RE Type using discourse status and acquisition group
as predictors. We used the brms package (Bürkner, 2018) in R, with a
categorical distribution on the responses and sum contrasts for the
predictors Discourse Status (Introduction, Maintenance, Re-introduction)
and Participant Group (Native, Late). As the overall number of produced
pronouns was very little (40 data points), we excluded pronominal use
from the following analyses. For the current analysis, we did not define
priors ourselves and used the preset uninformed priors. Figure
2\footnote{We ran another model where we used the Nativeness predictor
  with 3 levels 0-3:Native, 4-7:Late1, and 7-above:Late2. However, our
  model did not converge properly and the credible intervals were very
  wide spread, yielding the results uninformative. We provide the model
  results as a plot in \ref{section:d}.} shows the posterior probability
distributions where the dot is the median estimate, the thick line is
the 50\% credible intervals, and the thin line is the 95\% credible
intervals. There are 3 responses we are considering in this model. They
are Classifier, Nominative, and Null. Classifier response is the level
that other responses Nominative and Null are compared against.

@@

As for the distribution of the null markers, approximately 90\% of the
plain verbs were used to maintain a previous referent, making these
constructions least implicit REs used for TİD\footnote{As in other sign
  languages, in TİD, the lexical meaning of signs has an effect on the
  morphological grouping. For instance, usually (di)transitive agreement
  verbs denote some sort of physical or metaphorical transfer of an
  object or abstract notion (Meir, 1999). As for plain verbs, both
  phonology of a sign and its semantics have an effect on morphology.
  For instance, \textsc{eat} in TİD is signed nearby the mouth and since
  it needs to remain in that phonological location, it is not suitable
  for agreement and thus it is regarded as a plain verb. Likewise the
  experiencer verb \textsc{love} is signed on the torso and as such it
  does not take part in agreement. These generalizations noted we should
  also be cautious in labelling a verb as categorially one type or the
  other. For instance, some intransitive plain verbs like \textsc{fall}
  or \textsc{jump} in TİD may display spatial person agreement with the
  single argument in the clause (Gökgöz and Sevgi, 2020). See Lourenço
  and Wilbur (2018) also for a similar discussion in Brazilian Sign
  Language.\}.}. Agreement verbs and CA were the following implicit
markers of reference, 0.77 and 0.72 of which were used for maintained
contexts, respectively. This contrasts with some of the research
(Czubek, 2017) in which agreement verbs ranked lowest on scale,
indicating that these constructions were more implicit compared to plain
verbs. However, there seems to be a common ground for the position of CA
markers which have been claimed to rank higher than other null markers
as in the present study (Barberà and Massó, 2009; Czubek, 2017;
Frederiksen and Mayberry, 2016). In Figure \ref{fig:scale}, we report
the distribution of REs in TİD as a function of referent accessibility
and compare our accessibility hierarchy with that of (Frederiksen and
Mayberry, 2016) for ASL. In the hierarchies, the left most REs (e.g.,
modified and bare nouns for TİD) are used when the accessibility is low.
If the referent is already active in the addressee's mind and its
accessibility is high, then the right most markers (e.g., agreement and
plain verbs for TİD) are selected.

@@

\begin{figure}
\centering
\includegraphics{revised_manuscript_files/figure-latex/model-camera-1.pdf}
\caption{Regression model plot for the RE Type (Nominal, Null) with the
predictors Discourse Status (Introduction, Maintenance,
Re-introduction), Participant Group (Native, Late) and Narration Style
(all responses vs.~only camera)}
\end{figure}

\emph{Note:} ``All'' indicates model results including all participants
(both camera and experimenter) whereas ``camera'' indicates the results
from only those who narrated to the camera. That is, the 4 signers who
narrated to the experimenter were from analysis in the second dataset.

@@

\leavevmode\vadjust pre{\hypertarget{ref-gokgoz2020}{}}%
Gökgöz, Kadir, Sevgi, Hande, 2020. Aspects of clause structure and
morphology in turkish sign language. Morphological Complexity within and
across Boundaries: In Honour of Asli Göksel 215, 315.

@@

\leavevmode\vadjust pre{\hypertarget{ref-lourenuxe7o2018}{}}%
Lourenço, Guilherme, Wilbur, Ronnie B., 2018. Are plain verbs really
plain?: Co-localization as the agreement marker in sign languages.
FEAST. Formal and Experimental Advances in Sign language Theory 2, 6881.

@@

\leavevmode\vadjust pre{\hypertarget{ref-meir1999}{}}%
Meir, Irit, 1999. A perfect marker in israeli sign language. Sign
Language \& Linguistics 2, 4362.
